|

### 2023-06-19 paper.Rmd
small edits.
and revert back to using bookdown::pdf_document2 for cross-references.

---
title: |
  |
  | 
  | \vspace{1cm}Writing a reproducible paper
  | in R Markdown^[Corresponding address: mail@paulcbauer.de. Acknowledgments: We are grateful to all those generous people that invest their time into open-source software.]\vspace{0.5cm}
  |
author:
  - Paul Bauer, Mannheim Centre for European Social Research
  - Camille Landesvatter, Mannheim Centre for European Social Research
date: |
  |       
  | First version: December 14, 2018
  | This version: `r gsub("^0", "", format(Sys.time(), "%B %d, %Y"))`
  | Download: https://osf.io/q395s
  |
  | Feedback? Errors? mail@paulcbauer.de
  | [Github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)
  |
  | Tired of Latex?
  | Try [Pagedown](https://github.com/rstudio/pagedown) with our [new template](https://osf.io/k8jhx/)! :-)
linestretch: 1.2
colorlinks: true
abstract: \noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be downloaded [here](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown) (click on Code -> Download ZIP). We think we got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). We intend to update this file when we discover more convenient code (you can follow any updates through the corresponding [Github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).\vspace{.8cm}
bibliography: references.bib
csl: american-sociological-association.csl
output:
  pdf_document:
    toc: no
    keep_tex: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin=1in
always_allow_html: yes
---

@@

* **Access**: Research is normally funded by taxpayers (researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software. Importantly, researchers from developing countries are even more dependent on free access to knowledge [@Kirsop2005-ro].
* **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months. A fully reproducible setup will help you to trace back your own steps. Obviously, the same is true for other researchers who may want to understand your work and built on it. It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years. 
* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.
* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.rmd` file. For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.

# Prerequesites
We assume that you are using R on a day-to-day basis. You may have even started to work a little in R Markdown but you don't write your complete paper in R Markdown. If you don't know what R Markdown is watch [this short video](https://vimeo.com/178485416). Then...

@@

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(stargazer)
stargazer(cars, 
          title = "Summary table with stargazer",
          label="tab1", 
          table.placement = "H", 
          header=FALSE)
```

@@

## R base graphs
Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.rmd` file. It's relatively simple to insert R base graphs as you can see in Figure \@ref(fig:fig-1).

@@

```{r fig-2, fig.align="center", fig.cap="Miles per gallon according to the weight", fig.pos="H", fig.width=6, fig.height=3, message=FALSE, warning=FALSE, paged.print=FALSE}
mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
library(ggplot2)
ggplot(mtcars, aes(x=wt, y=mpg, shape=cyl)) + geom_point() +
  labs(x="Weight (lb/1000)", y = "Miles/(US) gallon", 
       shape="Number of \n Cylinders") + theme_classic()
```

>>>>

---
title: |
  | 
  | \vspace{0.7cm}Writing a reproducible paper
  | in R Markdown^[Corresponding address: mail@paulcbauer.de. Acknowledgments: We are grateful to all those generous people that invest their time into open-source software.]\vspace{0.5cm}
  |
author:
  - Paul Bauer, Mannheim Centre for European Social Research
  - Camille Landesvatter, Mannheim Centre for European Social Research
date: |
  |       
  | First version: December 14, 2018
  | This version: `r gsub("^0", "", format(Sys.time(), "%B %d, %Y"))`
  | Download this Paper on OSF: https://osf.io/q395s
  |
  | Feedback? Errors? mail@paulcbauer.de
  | [Github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)
  |
  | Tired of Latex?
  | Try [Pagedown](https://github.com/rstudio/pagedown) with our [new template](https://osf.io/k8jhx/)! :-)
linestretch: 1.2
colorlinks: true
abstract: \noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below, we outline some of the "tricks"/code (e.g., referencing tables, sections, etc.) we had to figure out to produce this document. The underlying files which produce this document can be downloaded [here](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown) (click on Code -> Download ZIP). We think we got pretty far, but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (e.g., bookdown). We intend to update this file when we discover more convenient code and you can follow any updates through the corresponding [Github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/).\vspace{.8cm}
bibliography: references.bib
csl: american-sociological-association.csl
output:
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin=1in
always_allow_html: yes
---

@@

* **Access**: Research is typically funded by taxpayers (and researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, such as the requirement of commercial software. Importantly, researchers from developing countries are particularly dependent on free access to knowledge [@Kirsop2005-ro].
* **Reproducibility**: Even if you have conducted a study and analyzed the data yourself, you may forget the steps taken after a few months. A fully reproducible setup will help you retrace your own steps. Additionally, other researchers who wish to understand and build upon your work will benefit from reproducibility. It may sound like a joke, but why not aim for a document that can be used to reproduce your findings even after 500 years?
* **Errors**: Manual steps in data analysis, such as manually copying/pasting values into a table, can introduce errors. R Markdown allows you to automate such steps and/or avoid them.
* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one .rmd file. For instance, if you decide to exclude a subset of your data, you simply need to insert one line of code at the beginning, and everything is automatically rebuilt/re-estimated.

# Prerequisites
We assume that you are using R on a day-to-day basis. You may have even started to work a little in R Markdown, but you don't write your complete paper in R Markdown. If you don't know what R Markdown is, watch [this short video](https://vimeo.com/178485416). Then...

@@

```{r tab1, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(stargazer)
stargazer(cars, 
          title = "Summary table with stargazer",
          label="tab1", 
          table.placement = "H", 
          header=FALSE)
```

@@

## R base graphs

Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.rmd` file. It's relatively simple to insert R base graphs as you can see in Figure \@ref(fig:fig-1))

@@

```{r fig-2, fig.align="center", fig.cap="Miles per gallon according to the weight", fig.pos="H", fig.width=6, fig.height=3, message=FALSE, warning=FALSE, paged.print=FALSE, echo=F}
mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor
library(ggplot2)
ggplot(mtcars, aes(x=wt, y=mpg, shape=cyl)) + geom_point() +
  labs(x="Weight (lb/1000)", y = "Miles/(US) gallon", 
       shape="Number of \n Cylinders") + theme_classic()
```

### 2023-06-19 paper.tex
small edits.
and revert back to using bookdown::pdf_document2 for cross-references.

\title{\hfill\break
\hfill\break
\vspace{1cm}Writing a reproducible paper\\
in R Markdown\footnote{Corresponding address: \href{mailto:mail@paulcbauer.de}{\nolinkurl{mail@paulcbauer.de}}. Acknowledgments: We are grateful to all those generous people that invest their time into open-source software.}\vspace{0.5cm}\\}
\author{Paul Bauer, Mannheim Centre for European Social Research \and Camille Landesvatter, Mannheim Centre for European Social Research}
\date{\hfill\break
First version: December 14, 2018\\
This version: April 05, 2023\\
Download: \url{https://osf.io/q395s}\\
\strut \\
Feedback? Errors? \href{mailto:mail@paulcbauer.de}{\nolinkurl{mail@paulcbauer.de}}\\
\href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/}{Github repository}\\
\strut \\
Tired of Latex?\\
Try \href{https://github.com/rstudio/pagedown}{Pagedown} with our \href{https://osf.io/k8jhx/}{new template}! :-)}

\begin{document}
\maketitle
\begin{abstract}
\noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below we outline some of the ``tricks''/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be downloaded \href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown}{here} (click on Code -\textgreater{} Download ZIP). We think we got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). We intend to update this file when we discover more convenient code (you can follow any updates through the corresponding \href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/}{Github repo}).\vspace{.8cm}
\end{abstract}

@@

\begin{itemize}
\tightlist
\item
  \textbf{Access}: Research is normally funded by taxpayers (researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software. Importantly, researchers from developing countries are even more dependent on free access to knowledge (\protect\hyperlink{ref-Kirsop2005-ro}{Kirsop and Chan 2005}).
\item
  \textbf{Reproducability}: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months. A fully reproducible setup will help you to trace back your own steps. Obviously, the same is true for other researchers who may want to understand your work and built on it. It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years.
\item
  \textbf{Errors}: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to \textbf{automatize} such steps and/or avoid them.
\item
  \textbf{Revisions}: Revising a paper takes much less time if you have all the code you need in one place, i.e., one \texttt{.rmd} file. For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.
\end{itemize}

\hypertarget{prerequesites}{%
\section{Prerequesites}\label{prerequesites}}

We assume that you are using R on a day-to-day basis. You may have even started to work a little in R Markdown but you don't write your complete paper in R Markdown. If you don't know what R Markdown is watch \href{https://vimeo.com/178485416}{this short video}. Then\ldots{}

@@

Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the \texttt{.rmd} file. It's relatively simple to insert R base graphs as you can see in Figure \ref{fig:fig-1}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mtcars}\SpecialCharTok{$}\NormalTok{cyl }\OtherTok{\textless{}{-}} \FunctionTok{as.factor}\NormalTok{(mtcars}\SpecialCharTok{$}\NormalTok{cyl) }\CommentTok{\# Convert cyl to factor}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{ggplot}\NormalTok{(mtcars, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x=}\NormalTok{wt, }\AttributeTok{y=}\NormalTok{mpg, }\AttributeTok{shape=}\NormalTok{cyl)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x=}\StringTok{"Weight (lb/1000)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Miles/(US) gallon"}\NormalTok{, }
       \AttributeTok{shape=}\StringTok{"Number of }\SpecialCharTok{\textbackslash{}n}\StringTok{ Cylinders"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{theme\_classic}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
## R version 4.2.2 (2022-10-31)
## Platform: aarch64-apple-darwin20 (64-bit)
## Running under: macOS Ventura 13.0.1
## 
## Matrix products: default
## BLAS:   /Library/Frameworks/R.framework/Versions/4.2-arm64/Resources/lib/libRblas.0.dylib
## LAPACK: /Library/Frameworks/R.framework/Versions/4.2-arm64/Resources/lib/libRlapack.dylib
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] ggplot2_3.4.2    kableExtra_1.3.4 knitr_1.42       stargazer_5.2.3 
## 
## loaded via a namespace (and not attached):
##  [1] compiler_4.2.2    pillar_1.9.0      tools_4.2.2       digest_0.6.31    
##  [5] evaluate_0.20     lifecycle_1.0.3   tibble_3.2.1      gtable_0.3.3     
##  [9] viridisLite_0.4.1 pkgconfig_2.0.3   rlang_1.1.0       cli_3.6.1        
## [13] rstudioapi_0.14   yaml_2.3.7        xfun_0.38         fastmap_1.1.1    
## [17] withr_2.5.0       dplyr_1.1.1       httr_1.4.5        stringr_1.5.0    
## [21] xml2_1.3.3        generics_0.1.3    vctrs_0.6.1       systemfonts_1.0.4
## [25] tidyselect_1.2.0  webshot_0.5.4     grid_4.2.2        svglite_2.1.1    
## [29] glue_1.6.2        R6_2.5.1          fansi_1.0.4       rmarkdown_2.21   
## [33] bookdown_0.33     farver_2.1.1      magrittr_2.0.3    scales_1.2.1     
## [37] htmltools_0.5.5   rvest_1.0.3       colorspace_2.1-0  labeling_0.4.2   
## [41] utf8_1.2.3        stringi_1.7.12    munsell_0.5.0
\end{verbatim}

>>>>

\title{\hfill\break
\vspace{0.7cm}Writing a reproducible paper\\
in R Markdown\footnote{Corresponding address: \href{mailto:mail@paulcbauer.de}{\nolinkurl{mail@paulcbauer.de}}. Acknowledgments: We are grateful to all those generous people that invest their time into open-source software.}\vspace{0.5cm}\\}
\author{Paul Bauer, Mannheim Centre for European Social Research \and Camille Landesvatter, Mannheim Centre for European Social Research}
\date{\hfill\break
First version: December 14, 2018\\
This version: June 19, 2023\\
Download this Paper on OSF: \url{https://osf.io/q395s}\\
\strut \\
Feedback? Errors? \href{mailto:mail@paulcbauer.de}{\nolinkurl{mail@paulcbauer.de}}\\
\href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/}{Github repository}\\
\strut \\
Tired of Latex?\\
Try \href{https://github.com/rstudio/pagedown}{Pagedown} with our \href{https://osf.io/k8jhx/}{new template}! :-)}

\begin{document}
\maketitle
\begin{abstract}
\noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below, we outline some of the ``tricks''/code (e.g., referencing tables, sections, etc.) we had to figure out to produce this document. The underlying files which produce this document can be downloaded \href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown}{here} (click on Code -\textgreater{} Download ZIP). We think we got pretty far, but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (e.g., bookdown). We intend to update this file when we discover more convenient code and you can follow any updates through the corresponding \href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/}{Github repo}.\vspace{.8cm}
\end{abstract}

@@

\begin{itemize}
\tightlist
\item
  \textbf{Access}: Research is typically funded by taxpayers (and researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, such as the requirement of commercial software. Importantly, researchers from developing countries are particularly dependent on free access to knowledge (\protect\hyperlink{ref-Kirsop2005-ro}{Kirsop and Chan 2005}).
\item
  \textbf{Reproducibility}: Even if you have conducted a study and analyzed the data yourself, you may forget the steps taken after a few months. A fully reproducible setup will help you retrace your own steps. Additionally, other researchers who wish to understand and build upon your work will benefit from reproducibility. It may sound like a joke, but why not aim for a document that can be used to reproduce your findings even after 500 years?
\item
  \textbf{Errors}: Manual steps in data analysis, such as manually copying/pasting values into a table, can introduce errors. R Markdown allows you to automate such steps and/or avoid them.
\item
  \textbf{Revisions}: Revising a paper takes much less time if you have all the code you need in one place, i.e., one .rmd file. For instance, if you decide to exclude a subset of your data, you simply need to insert one line of code at the beginning, and everything is automatically rebuilt/re-estimated.
\end{itemize}

\hypertarget{prerequisites}{%
\section{Prerequisites}\label{prerequisites}}

We assume that you are using R on a day-to-day basis. You may have even started to work a little in R Markdown, but you don't write your complete paper in R Markdown. If you don't know what R Markdown is, watch \href{https://vimeo.com/178485416}{this short video}. Then\ldots{}

@@

Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the \texttt{.rmd} file. It's relatively simple to insert R base graphs as you can see in Figure \ref{fig:fig-1})

@@

@@

\begin{verbatim}
## R version 4.2.2 (2022-10-31)
## Platform: aarch64-apple-darwin20 (64-bit)
## Running under: macOS Ventura 13.2.1
## 
## Matrix products: default
## BLAS:   /Library/Frameworks/R.framework/Versions/4.2-arm64/Resources/lib/libRblas.0.dylib
## LAPACK: /Library/Frameworks/R.framework/Versions/4.2-arm64/Resources/lib/libRlapack.dylib
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] ggplot2_3.4.2         kableExtra_1.3.4.9000 knitr_1.42           
## [4] stargazer_5.2.3      
## 
## loaded via a namespace (and not attached):
##  [1] compiler_4.2.2    pillar_1.9.0      tools_4.2.2       digest_0.6.31    
##  [5] evaluate_0.20     lifecycle_1.0.3   tibble_3.2.1      gtable_0.3.3     
##  [9] viridisLite_0.4.1 pkgconfig_2.0.3   rlang_1.1.0       cli_3.6.1        
## [13] rstudioapi_0.14   yaml_2.3.7        xfun_0.38         fastmap_1.1.1    
## [17] withr_2.5.0       dplyr_1.1.1       httr_1.4.5        stringr_1.5.0    
## [21] xml2_1.3.3        generics_0.1.3    vctrs_0.6.1       systemfonts_1.0.4
## [25] tidyselect_1.2.0  webshot_0.5.4     grid_4.2.2        svglite_2.1.1    
## [29] glue_1.6.2        R6_2.5.1          fansi_1.0.4       rmarkdown_2.21   
## [33] bookdown_0.33     farver_2.1.1      magrittr_2.0.3    scales_1.2.1     
## [37] htmltools_0.5.5   rvest_1.0.3       colorspace_2.1-0  labeling_0.4.2   
## [41] utf8_1.2.3        stringi_1.7.12    munsell_0.5.0
\end{verbatim}
